---
title: "AggregateData"
author: "Ben Weinstein"
date: "May 2, 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

Aim: To create a generalized cleaning structure for the Ecuador data.

The date structures remain a challenge.

```{r setup, include=FALSE}
library(maptools)
library(plotKML)
library(ggplot2)
library(taxize)
library(dplyr)
library(stringr)
library(stringi)
library(chron)
library(readxl)
library(taxize)
library(anytime)
library(googlesheets)
basename="/Users/Ben/Dropbox/HummingbirdProject/Data/"
```

# Read taxonomy

From Fransisco's dropbox. Live updating to reflect new identifications.

```{r,eval=T}
taxsheet<-read_excel("/Users/ben/Dropbox/Lista Planta final/Lista planta final_12082019.xlsx") %>% select(Site,"field_name (AyC)","ID_final") %>% rename(plant_field_name=`field_name (AyC)`,final_plant_name=`ID_final`)

abb<-data.frame(Site=c("RUMI","MSPL","VECO","MSPC","ALAS","GRAL","UPDC","MAQUI","UPDCD","YANA","YANAD","ALASD","STLU","PUYU","SACH","STLL","SACHD","AMAG","SANTA","AMAGD","BELLA","CAMBA"),FullSite=c("Rumisitana","MashpiLaguna","Verdecocha","MashpiCapuchin","Alaspungo","LasGralarias","UnPocoDelChoco","Maquipucuna","UnPocoDelChoco_disturbed","Yanacocha","Yanacocha_disturbed","Alaspungo_disturbed","SantaLuciaUpper","Puyucunapi","Sachatamia","SantaLuciaLower","Sachatamia_disturbed","Amagusa","Santa Lucia","Amagusa_disturbed","Bellavista","Cambugan"))
taxsheet<-taxsheet %>% left_join(abb)

#Do some cap standards, only the first word is capitalized
taxsheet$plant_field_name<-tolower(taxsheet$plant_field_name)
taxsheet$plant_field_name<-stri_trans_totitle(taxsheet$plant_field_name,opts_brkiter = stri_opts_brkiter(type = "sentence"))

taxsheet$final_plant_name<-tolower(taxsheet$final_plant_name)
taxsheet$final_plant_name<-stri_trans_totitle(taxsheet$final_plant_name,opts_brkiter = stri_opts_brkiter(type = "sentence"))

#remove all periods
taxsheet$plant_field_name<-str_replace(taxsheet$plant_field_name,"\\.","")
taxsheet$final_plant_name<-str_replace(taxsheet$final_plant_name,"\\.","")

#no space between sp and number
taxsheet$plant_field_name<-str_replace(taxsheet$plant_field_name,"sp (\\d)","sp\\1")
taxsheet$final_plant_name<-str_replace(taxsheet$final_plant_name,"sp (\\d)","sp\\1")

```

# GPS Data

Read gpx files uploaded by field researchers.

```{r}
allwaypoints<-list.files(basename,recursive=TRUE,full.name=TRUE,pattern=".gpx")
```

```{r}
#read in waypoints
wayfiles<-function(x){
    try(k<-readGPX(x,waypoints=TRUE)$waypoints)
    if(!exists("k")){
      print("missing")
      return(NULL)
    }
    k$site<-str_match(x,"(\\w+)/waypoints")[2]
    if("cmt" %in% (colnames(k))){
      k$time<-k$cmt
    }
    
    #remove some generic files
    return(k)
  }
wpoints<-lapply(allwaypoints,wayfiles)
wpoints<-wpoints[!sapply(wpoints,length)==1]

wpoints<-rbind_all(wpoints)

#strip those points with no time or elevation
missing_elev<-wpoints %>% filter(is.na(ele))
write.csv(missing_elev,"MissingElevation.csv")

wpoints<-wpoints %>% filter(!is.na(ele))
wpoints<-wpoints %>% filter(!is.na(time))

wpoints$waypoint<-wpoints$name

#Give it a try
wpoints$timestamp<-anytime(wpoints$time)
missing<-wpoints %>% filter(is.na(timestamp)) %>% select(time) %>% .$time

for(x in missing){
  wpoints$timestamp[wpoints$time %in% x]<-as.POSIXct(x,format="%d-%b-%y %H:%M:%S")
}

#look again
missing<-wpoints %>% filter(is.na(timestamp)) %>% select(time) %>% .$time

#some missing april data in spanish
for(x in missing){
  wpoints$timestamp[wpoints$time %in% x]<-as.POSIXct(str_replace(x,"ABR","APR"),format="%d-%b-%y %H:%M:%S")
}

#look again
missing<-wpoints %>% filter(is.na(timestamp)) %>% select(time) %>% .$time

#That looks sufficient for covering dates.
wpoints<-wpoints %>% filter(!is.na(timestamp))

#turn to numeric if has a leading 0, 006 -> 6
wpoints$waypoint[substring(wpoints$waypoint,1,1)=="0"]<-as.numeric(wpoints$waypoint[substring(wpoints$waypoint,1,1)=="0"])

#Create ids
#create site id, month, year waypoint combination.
wpoints$Month<-months(wpoints$timestamp)

#delete duplicates and unneeded columns
wpoints<-wpoints[!duplicated(wpoints),]

wpoints<-wpoints %>% select(-sym,-type,-extensions,-link,-cmt,-desc,-name)

#if there are multiple points per waypoint, take the latest in time.
return_latest<-function(x){
  df<-x %>% arrange(desc(time)) %>% head(1)
  return(df)
}
wpoints<-wpoints %>% group_by(site,Month,waypoint) %>% do(return_latest(.))
```

```{r}
#write gps points
write.csv(wpoints,"/Users/Ben/Dropbox/HummingbirdProject/Data/HummingbirdProjectCleaned/GPS.csv")

#Write in case we want to visualize with the shiny data
write.csv(wpoints,"HummingbirdData/GPS.csv")
```

# Transect Data

```{r}
transect_files<-list.files(basename,recursive=TRUE,pattern="transect_",full.names=T)

transect_files<-transect_files[!str_detect(transect_files,"~")]
transect_files<-transect_files[!str_detect(transect_files,"metadata")]

#bind site name by folder

transect_xlsx<-lapply(transect_files,function(x){
  
  print(x)
  if(str_detect(x,".xlsx")){
    y<-read_xlsx(x)
    
    #if empty, just move to next file
    if(nrow(y)==0){
      return(y)
    }
    
    y$site<-str_match(x,"(\\w+)/transect")[2]

    #turn waypoints to character for the moment
    y$waypoint<-as.character(y$waypoint)
    y$total_flowers<-as.character(y$total_flowers)
    y$height<-as.character(y$height)
    y$count_method<-as.character(y$count_method)
    y$flower_unit<-as.character(y$flower_unit)
    y$flower_count1<-as.character(y$flower_count1)
    y$flower_count3<-as.character(y$flower_count3)
    y$date<-as.character(y$date)

  }
  if(str_detect(x,".csv")){
    
    y<-read.csv(x)
    y$site<-str_match(x,"(\\w+)/transect")[2]

    #turn waypoints to character for the moment
    y$height<-as.character(y$height)
    y$waypoint<-as.character(y$waypoint)
    y$total_flowers<-as.character(y$total_flowers)
    y$flower_unit<-as.character(y$flower_unit)
    y$flower_count1<-as.character(y$flower_count1)
    y$flower_count3<-as.character(y$flower_count3)
    y$count_method<-as.character(y$count_method)

  }
  
  return(y)
})

#remove empty files
transect_xlsx<-transect_xlsx[!sapply(transect_xlsx,nrow)==0]
#Combine files
transect_data<-bind_rows(transect_xlsx)

#turn total flowers back to numeric
transect_data$total_flowers<-as.numeric(transect_data$total_flowers)

#replace missing cells with NA, todo check holger's ugly file format
transect_data$revised_plant_name[transect_data$revised_plant_name==""]<-NA

#If revised plant name exists, replace
transect_data$taxize_plant_name<-transect_data$plant_field_name
for(x in 1:nrow(transect_data)){
  if(!is.na(transect_data$revised_plant_name[x])){
    transect_data$taxize_plant_name[x]<-transect_data$revised_plant_name[x]
  }
}

#Do some cap standards, only the first word is capitalized
transect_data$taxize_plant_name<-tolower(transect_data$taxize_plant_name)
transect_data$taxize_plant_name<-stri_trans_totitle(transect_data$taxize_plant_name,opts_brkiter = stri_opts_brkiter(type = "sentence"))

#remove all periods
transect_data$taxize_plant_name<-str_replace(transect_data$taxize_plant_name,"\\.","")

#no space between sp and number
transect_data$taxize_plant_name<-str_replace(transect_data$taxize_plant_name,"sp (\\d)","sp\\1")
```

## Taxonomy Cleaning

This comes in three steps.

* Basic name standardization. First word is capitalized, second is not. Drop special characters and carriage returns.

* Spelling mistakes are checked by looking at the Tropicos database using taxize. Recognized names are replaced, the rest are left in place.

* Check against fransisco's identification sheet. 

```{r}
#Month and year column
transect_data$Month<-months(strptime(transect_data$date,"%d/%m/%Y"))
transect_data$Year<-years(strptime(transect_data$date,"%d/%m/%Y"))

#Look up taxize names against tropicos
sources <- gnr_datasources()
tropicos<-sources$id[sources$title == 'Tropicos - Missouri Botanical Garden']

#Create plant taxonomy factors by site (required for fransisco's data)
tran_levels<-transect_data %>% distinct(taxize_plant_name) %>% filter(!is.na(taxize_plant_name))

#send to tropicos
tran_taxize<-gnr_resolve(tran_levels$taxize_plant_name,best_match_only=T,canonical = TRUE,data_source_ids=tropicos) 

#only keep species with matched double names - the rest are unknown to tropicos
plants_keep<-tran_taxize %>% filter(matched_name2 %in% tran_taxize$matched_name2[str_count(tran_taxize$matched_name2,"\\w+")==2]) %>% .$user_supplied_name

#replace with taxized names, fixes spelling errors
for(plant_name in plants_keep){
  transect_data[transect_data$taxize_plant_name %in% plant_name,"taxize_plant_name"]<-tran_taxize[tran_taxize$user_supplied_name == plant_name,"matched_name2"]
}
```

```{r}
#matched with fransiscos names based on site. Drop unknown for now. Not convinced this is complete. Cannot join, need to proceed slowly.
transect_data$final_plant_name<-transect_data$taxize_plant_name
for(i in 1:nrow(taxsheet)){
  transect_data[transect_data$site %in% taxsheet$site[i] & transect_data$taxize_plant_name %in% taxsheet$taxize_plant_name[i],"final_plant_name"]<-taxsheet$final_plant_name[i]
}

#Missing taxonomy defined as neither in the tropicos or fransisco's taxonomy database
final_levels<-unique(transect_data$final_plant_name)
not_in_tropicos<-final_levels[!final_levels %in% tran_taxize$matched_name2]
plants_missing<-not_in_tropicos[!not_in_tropicos %in% taxsheet$final_plant_name]
```

What percentage of transect names remain missing?

```{r}
transect_data %>% filter(!is.na(final_plant_name),!final_plant_name %in% c(" ","")) %>% mutate(missing=final_plant_name %in% plants_missing) %>% group_by(site) %>% summarize(n=sum(missing)/n()*100)
```

```{r}
#Create missing plant index.
plant_tax_needed<-transect_data %>% filter(final_plant_name %in% plants_missing,!is.na(final_plant_name),!final_plant_name %in% c(" ",""))  %>% group_by(site) %>% distinct(final_plant_name)
write.csv(plant_tax_needed,"missing_taxonomy.csv")

#write for each site
sites<-unique(transect_data$site)
for(x in 1:length(sites)){
 plant_tax_needed %>% filter(site==sites[x]) %>% write.csv(.,paste("/Users/Ben/Dropbox/HummingbirdProject/Data/",sites[x],"Transects_Missing_Taxonomy.csv",sep="/"))
}
```

```{r}
transect_data<-transect_data %>% filter(!final_plant_name %in% plants_missing) %>% 
select(date,Month,Year,site,final_plant_name,total_flowers,waypoint,height,hummingbird,comment) %>% droplevels()
```

# Clean hummingbird names

```{r}
#get species list
hum_levels<-levels(transect_data$hummingbird)

#lookup taxize
hum_taxize<-gnr_resolve(hum_levels,best_match_only=T,canonical=TRUE)

#only keep species with matched double names
hum_keep<-hum_taxize$submitted_name[sapply(strsplit(hum_taxize$matched_name2," "),length)==2]
hum_missing<-hum_taxize$submitted_name[!sapply(strsplit(hum_taxize$matched_name2," "),length)==2]

#for formatted lavels
for (x in levels(transect_data$hummingbird)){
  levels(transect_data$hummingbird)[levels(transect_data$hummingbird) %in% x]<-hum_taxize %>% filter(submitted_name %in% x) %>% .$matched_name2  
}
```

## Combine transect with gps data

```{r}
transect_gps<-transect_data %>% left_join(wpoints,by=c("site","Month","waypoint"))

paste(missing<-transect_gps %>% filter(is.na(lon)) %>% nrow(.), "transect points", "out of", nrow(transect_gps), "missing GPS data")
missing_transect_gps<-transect_gps %>% filter(is.na(lon)) %>% group_by(site,date) %>% summarize(n=n()) %>% arrange(desc(n))

write.csv(missing_transect_gps,"missing_transect_dates.csv")
```

Generate error scripts

```{r}
for(x in 1:length(sites)){
  missing_transect_gps %>% filter(site==sites[x]) %>% write.csv(.,paste("/Users/Ben/Dropbox/HummingbirdProject/Data/",sites[x],"Transects_Missing_GPS.csv",sep="/"))
}
```

## Divide into plant and hummingbird transects

This is producing empty hummingbird transect data!

```{r}
hummingbird_transects<-transect_gps %>% filter(!hummingbird=="")

#only hummingbirds
Sys.setenv(ENTREZ_KEY="f00b6e4c6b153028dfbe9389cd18f0a46b08")
hummingbird_transects$hummingbird<-as.factor(hummingbird_transects$hummingbird)
isclass<-tax_name(query = levels(hummingbird_transects$hummingbird), get = "family", db="ncbi")
troch_keep<-isclass %>% filter(family %in% c("Trochilidae","Thraupidae")) %>% .$query
hummingbird_transects<-hummingbird_transects %>% filter(hummingbird %in% troch_keep) %>% droplevels()

#What's up with cambugan, alot of no total flower counts?
plant_transects<-transect_gps %>% filter(!final_plant_name=="") %>% filter(!is.na(total_flowers))
```

## Write to file

```{r}
#plants
write.csv(hummingbird_transects,"/Users/Ben/Dropbox/HummingbirdProject/Data/HummingbirdProjectCleaned/HummingbirdTransects.csv")

#write to shiny server
write.csv(hummingbird_transects,"HummingbirdData/HummingbirdTransects.csv")

#plants
write.csv(plant_transects,"/Users/Ben/Dropbox/HummingbirdProject/Data/HummingbirdProjectCleaned/PlantTransects.csv")

#write to shiny server
write.csv(plant_transects,"HummingbirdData/PlantTransects.csv")
```

# Camera Data

```{r}
camera_files<-list.files(basename,recursive=TRUE,pattern="cameras",full.names = T)

camera_xlsx<-lapply(camera_files,function(x){
  print(x)
  if(str_detect(x,".xlsx")){
    
    j<-read_xlsx(x)
    #if empty, just move to next file
    if(nrow(j)==0){
      return(j)
    }
    
    #format columns
    j$site<-str_match(x,"(\\w+)/cameras")[2]
    j$waypoint<-as.character(j$waypoint)
    j$card_id<-as.character(j$card_id)
    j$start_time<-as.character(j$start_time)
    j$start_date<-as.character(j$start_date)
    j$end_date<-as.character(j$end_date)
    j$end_time<-as.character(j$end_date)
    j$height<-as.character(j$height)
    j$X__1<-NULL
    return(j)
  }
  
  if(str_detect(x,".csv")){
    j<-read.csv(x)
    j$site<-str_match(x,"(\\w+)/cameras")[2]
    j$waypoint<-as.character(j$waypoint)
    j$card_id<-as.character(j$card_id)
    j$start_time<-as.character(j$start_time)
    j$height<-as.character(j$height)

    return(j)
  }
  
})

camera_dat<-bind_rows(camera_xlsx)

#select desired columns
camera_dat<-camera_dat %>% select(start_date,start_time,end_date,end_time,site,waypoint,plant_field_name,revised_plant_name)

#Month and year column
camera_dat$Month<-months(strptime(camera_dat$start_date,"%d/%m/%Y"))
camera_dat$Year<-years(strptime(camera_dat$start_date,"%d/%m/%Y"))
```

## Taxonomy Cleaning

```{r}
#replace missing cells with NA, todo check holger's ugly file format
camera_dat$revised_plant_name[camera_dat$revised_plant_name==""]<-NA

#If revised plant name exists, replace
camera_dat$taxize_plant_name<-camera_dat$plant_field_name
for(x in 1:nrow(camera_dat)){
  if(!is.na(camera_dat$revised_plant_name[x])){
    camera_dat$taxize_plant_name[x]<-camera_dat$revised_plant_name[x]
  }
}

#Do some cap standards, only the first word is capitalized
camera_dat$taxize_plant_name<-tolower(camera_dat$taxize_plant_name)
camera_dat$taxize_plant_name<-stri_trans_totitle(camera_dat$taxize_plant_name,opts_brkiter = stri_opts_brkiter(type = "sentence"))

#remove all periods
camera_dat$taxize_plant_name<-str_replace(camera_dat$taxize_plant_name,"\\.","")

#no space between sp and number
camera_dat$taxize_plant_name<-str_replace(camera_dat$taxize_plant_name,"sp (\\d)","sp\\1")

#Look up taxize names against tropicos
sources <- gnr_datasources()
tropicos<-sources$id[sources$title == 'Tropicos - Missouri Botanical Garden']

#Create plant taxonomy factors by site (required for fransisco's data)
tran_levels<-camera_dat %>% distinct(taxize_plant_name) %>% filter(!is.na(taxize_plant_name))

#send to tropicos
tran_taxize<-gnr_resolve(tran_levels$taxize_plant_name,best_match_only=T,canonical = TRUE,data_source_ids=tropicos) 

#only keep species with matched double names - the rest are unknown to tropicos
plants_keep<-tran_taxize %>% filter(matched_name2 %in% tran_taxize$matched_name2[str_count(tran_taxize$matched_name2,"\\w+")==2]) %>% .$user_supplied_name

#replace with taxized names, fixes spelling errors
for(plant_name in plants_keep){
  camera_dat[camera_dat$taxize_plant_name %in% plant_name,"taxize_plant_name"]<-tran_taxize[tran_taxize$user_supplied_name == plant_name,"matched_name2"]
}
```

```{r}
#matched with fransiscos names based on site. Drop unknown for now. Not convinced this is complete. Cannot join, need to proceed slowly.
camera_dat$final_plant_name<-camera_dat$taxize_plant_name
for(i in 1:nrow(taxsheet)){
  camera_dat[camera_dat$site %in% taxsheet$site[i] & camera_dat$taxize_plant_name %in% taxsheet$taxize_plant_name[i],"final_plant_name"]<-taxsheet$final_plant_name[i]
}

#Missing taxonomy defined as neither in the tropicos or fransisco's taxonomy database
final_levels<-unique(camera_dat$final_plant_name)
not_in_tropicos<-final_levels[!final_levels %in% tran_taxize$matched_name2]
plants_missing<-not_in_tropicos[!not_in_tropicos %in% taxsheet$final_plant_name]
```

What percentage of cameras names remain missing?

```{r}
camera_dat %>% filter(!is.na(final_plant_name),!final_plant_name %in% c(" ","")) %>% mutate(missing=final_plant_name %in% plants_missing) %>% group_by(site) %>% summarize(n=sum(missing)/n()*100)
```

```{r}
#Create missing plant index.
plant_tax_needed<-camera_dat %>% filter(final_plant_name %in% plants_missing,!is.na(final_plant_name),!final_plant_name %in% c(" ","")) %>% distinct(site,final_plant_name)
write.csv(plant_tax_needed,"camera_missing_taxonomy.csv")

#write for each site
sites<-unique(transect_data$site)
for(x in 1:length(sites)){
 plant_tax_needed %>% filter(site==sites[x]) %>% write.csv(.,paste("/Users/Ben/Dropbox/HummingbirdProject/Data/",sites[x],"Cameras_Missing_Taxonomy.csv",sep="/"))
}
```

## Combine with gps data

```{r}
#make sure wpoints is singular with respect to site waypoints.
gps_tojoin_camera<-wpoints %>% filter(waypoint %in% camera_dat$waypoint ) %>% group_by(site,waypoint) %>% summarize(lon=mean(lon),lat=mean(lat),ele=mean(as.numeric(ele),na.rm=T),time=max(time),Month=max(Month))

camera_gps<-camera_dat %>% select(-Month) %>% left_join(gps_tojoin_camera,by=c("site","waypoint")) 
paste(camera_gps %>% filter(is.na(lon)) %>% nrow(.), "camera points missing GPS data from",  nrow(camera_gps))
camera_gps %>% filter(is.na(lon)) %>% group_by(site) %>% summarize(n=n()) %>% arrange(desc(n))
```

## Write to file

```{r}
#write to dropbox
write.csv(camera_dat,"/Users/Ben/Dropbox/HummingbirdProject/Data/HummingbirdProjectCleaned/Cameras.csv")

#write to shiny server
write.csv(camera_dat,"HummingbirdData/Cameras.csv")
```

# Interaction Data

```{r}
int_files<-list.files(basename,recursive=TRUE,pattern="observations",full.names = T)

#TODO confirm with nicole that this can be dropped
int_files<-int_files[!int_files %in% "/Users/Ben/Dropbox/HummingbirdProject/Data//UnPocoDelChoco/Disturbio/foundframes/Disturbio/observations_UPDC-Disturbio.xlsx"]

int_data<-lapply(int_files,function(x){
  if(str_detect(x,".xlsx")){
    j<-read_excel(x,trim_ws = T)
    try(j$site<-str_match(x,"(\\w+)/observations")[2],silent=T)
    j$time<-as.character(j$time)
    j$filename<-as.character(j$filename)
    j$date<-as.character(j$date)
    j$waypoint<-as.character(j$waypoint)
    return(j)
  }
  if(str_detect(x,".csv")){
    j<-read.csv(x)
    try(j$site<-str_match(x,"(\\w+)/observations")[2],silent=T)
    j$time<-as.character(j$time)
    j$filename<-as.character(j$filename)
    j$date<-as.character(j$date)
    j$waypoint<-as.character(j$waypoint)
    return(j)
  }
  
})

int_data<-bind_rows(int_data)

#drop duplicate rows
int_data<-int_data[!duplicated(int_data),]
```

## Cleaning

### Hummingbird names

The status of Phaethornis baroni needs to be fixed? Not clear if this is an accepted at species or subspecies level. Taxize is rejecting it. Thalurania colombica fannyi too.

```{r}
#conservatively remove questionable records
int_data<-int_data %>% filter(!is.na(hummingbird))
int_data$hummingbird[int_data$hummingbird %in% "Ocreatus underwoodi"]<-"Ocreatus underwoodii"

#remove ? from names, treated as wildcard for taxize
int_data<-int_data[!str_detect(int_data$hummingbird,"\\?"),] 
  
#refactor
int_data$hummingbird<-as.factor(int_data$hummingbird)

#get species list
int_levels<-levels(int_data$hummingbird)

#fill in revised names
int_data$revised_hummingbird[int_data$revised_hummingbird %in% c("h", " ","")]<-NA
int_data[!is.na(int_data$revised_hummingbird),"hummingbird"]<-int_data$revised_hummingbird[!is.na(int_data$revised_hummingbird)]
int_data[int_data$hummingbird %in% "","hummingbird"]<-NA
int_data$hummingbird<-factor(int_data$hummingbird)

#lookup taxize
int_taxize<-gnr_resolve(int_levels,best_match_only=T,canonical=TRUE)

#only keep species with matched double names
hum_keep<-int_taxize$submitted_name[sapply(strsplit(int_taxize$matched_name2," "),length)==2]
hum_missing<-int_taxize$submitted_name[!sapply(strsplit(int_taxize$matched_name2," "),length)==2]

int_data<-int_data %>% filter(hummingbird %in% hum_keep) %>% droplevels()

#for formatted lavels
for (x in levels(int_data$hummingbird)){
  levels(int_data$hummingbird)[levels(int_data$hummingbird) %in% x]<-int_taxize %>% filter(submitted_name %in% x) %>% .$matched_name2  
}

int_data<-droplevels(int_data)

#only hummingbirds
Sys.setenv(ENTREZ_KEY="f00b6e4c6b153028dfbe9389cd18f0a46b08")
isclass<-tax_name(query = levels(int_data$hummingbird), get = "family", db = "ncbi")
troch_keep<-isclass %>% filter(family %in% c("Trochilidae","Thraupidae")) %>% .$query
int_data<-int_data %>% filter(hummingbird %in% troch_keep) %>% droplevels()
```

```{r}
### Holger had one month where he entered all the frames.
#Loop through frames and remove consecutives.
diff_frame<-function(x){
  framenum<-as.numeric(str_match(x$filename,"(\\w+).jpg")[,2])
  out<-c()
  if(length(framenum)==1){
    out<-NA
    return(data.frame(ID=unique(x$waypoint),filename=NA,frames_since=out))
  }
    for(i in 2:length(framenum)){
      out[i]<-framenum[i]-framenum[i-1]
    }
  return(data.frame(ID=unique(x$waypoint)[1],filename=x$filename,frames_since=out))
}

int_data$ID<-1:nrow(int_data)
too_close<-int_data %>% group_by(folder) %>% do(diff_frame(.)) %>% filter(!is.na(frames_since),frames_since < 20) %>% rename(waypoint=ID)

int_data<-int_data %>% anti_join(too_close)

#fill empty dates
a<-int_data %>% filter(date=="") %>% mutate(newdate=str_match(folder,"/(\\d+)")[,2]) %>% .$newdate
int_data[int_data$date %in% "","date"]<-format(as.POSIXct(a,format="%y%m%d",tz=""),"%d/%m%/%Y")
```

## Combine with camera data

```{r}
tojoin<-camera_gps %>% select(site,waypoint,final_plant_name,lon,lat,ele)
int_gps<-int_data %>% left_join(tojoin,by=c("site","waypoint")) %>% select(ID,hummingbird,final_plant_name,folder,filename,date,time,revised_hummingbird,sex,piercing,site,waypoint,lon,lat,ele)

#TODO check replicates here
int_gps<-int_gps[!duplicated(int_gps),]
```

```{r}
#label events
int_gps$ID=1:nrow(int_gps)

#drop without hummingbird data
int_gps<-int_gps %>% filter(!is.na(hummingbird))

#missing gps data
paste(int_gps %>% filter(is.na(lon)) %>% nrow(.), "records missing gps data", "from",nrow(int_gps))

print("Missing GPS records by site")
int_gps %>% filter(is.na(lon)) %>% group_by(site) %>% summarize(n=n()) %>% arrange(desc(n))

print("Missing GPS cameras by site")
int_gps %>% filter(is.na(lon)) %>% group_by(site) %>% summarize(n=length(unique(waypoint))) %>% arrange(desc(n))
```

## Write to file

```{r}
#write to dropbox
write.csv(int_gps,"/Users/Ben/Dropbox/HummingbirdProject/Data/HummingbirdProjectCleaned/Interactions.csv")

#write to shiny server
write.csv(int_gps,"HummingbirdData/Interactions.csv")
```

